Probability of causal contact between interstellar civilizations through MonteCarlo simulations

### 2020-01-08 Sec_results.tex
retocando las conclusiones

\ffn{3}
%
In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
distributions of the number contacts for \cetis{} in four different
samples (M5, M6, M7, M8) with long range signal reach
(D$_{max}$=10000). 
%
The lifetime is more determinant than the awakening rate in increasing
the number of contacts.
%
As expected, of the four samples, the sample M7, which corresponds to
a dense awakening in the timeline and a long lifetime, is the model
that maximizes the number of contacts, reaching a maximum of order 100
contacts for a single \ceti{}.
%
This is considerable larger than any model with a shorter lifetime,
which produce a number of contacts of at most the order of ten
contacts per \ceti{}.
%
A couple of comments are worth to mention about this plot.
%
First, it has a logarithmic scale on the x-axis, and second it is the
cumulative, not differential, empirical distribution.

@@

%{{{
\ffn{4, 5}
%
We show the 2D maps of the rate of \cetis{} that never make contact and
the rate of \cetis{} that make contact at the moment of the awakening,
in the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
as a function of $\tau_s$ and $\tau_s$.
%
DISCUTIR LAS FIGURAS 4 Y 5 !!!

@@

In this section we analyze the distribution of the waiting times for
a first contact.
%
%Such distribution can be considered to compute the probability of
%reaching a first contact within the first XX years of a SETI project,
%assuming the methods are correct and the hypotheses of the experiment.
Such distribution can be considered to compute the probability for a
random \ceti{} to carry out a SETI project and spend a given time until
the first contact is made.

\ffn{6}
%
In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
waiting times for the first contact, for several models.
%
Left panel shows the histograms for several values of $\tau_a$, and
$\tau_s$ in the range 17000-24000 yr.
%
Right panel shows the histograms for several values of $\tau_s$, and
$\tau_a$ in the range 68000-100000 yr.
%
The shape of the histograms does not change when $\tau_a$ is modified,
but significantly changes when $\tau_s$ is modified.

In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the
\cetis{}
that have made at least one contact as a function of the elapsed time
since the awakening, given that they make at least one contact on
their lifetime, for different models.
%
From a frequentist approach, the normalized cumulative counts are
related to the estimation of the probability of observing a given time
with no success.
%             
Both panels show the histograms of the waiting times for the first
contact, for all the \cetis{} that make at least one contact along their
lifetime.
%
We show separately the histograms for several values of $\tau_a$ and a
fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
and a fixed value of $(\tau_a, D_{max})$.
%
For different values of $\tau_a$, the shape of the distribution is
similar, and the main difference given by the different number of
contacts on the different models.
%
Conversely, for different values of $\tau_a$, the shape of the
distribution is similar, and the main difference given by the
different number of contacts on the different models.

>>>>

@@

%{{{

\ffn{3}
%
In Fig.~\ref{F_number_of_contacts} we show the empirical cumulative
distributions of the number contacts for \cetis{} in eight different
samples, including short and long lifetimes
$\left(\,\left<\tau_s\right>=\right.$10000~yr and  $\left<\tau_s\right>$=100000~yr,
respectively$\left.\right)$, dense and sparse spatial distribution
$\left(\,\left<\tau_a\right>=\right.$10000~yr and $\left<\tau_a\right>$= 100000~yr,
respectively$\left.\right)$, and two different signal reach ranges (D$_{max}$=40000
lyr and D$_{max}$=80000~lyr). 
%
The areas between models with the same mean lifetime and mean
awakening time are shaded for visualization purposes.
%
As it can be seen from this Figure, the mean lifetime is more
determinant than the mean awakening rate for the number of contacts.
%
As expected, of the four samples with the longest $D_{max}$ the model
with a dense awakening in the timeline (low $\left<\tau_a\right>$) and
a long lifetime (high $\left<\tau_s\right>$), is the model that
maximizes the number of contacts, reaching a maximum of order 20
contacts for a single \ceti{}.
%
Similarly, for the models with long signal range
($D_{max}=80000\,lyr$), the one which is dense and with long lifetime
has the maximum number of contacts, reaching nearly 100 contacts for
each \ceti{} in its entire lifetime.
%
This is considerable larger than any model with a shorter lifetime,
which produce a number of contacts of at most the order of ten
contacts per \ceti{}.
%
A couple of comments are worth to mention about this plot.
%
First, it has a logarithmic scale on the x-axis, and second it is the
cumulative, not differential, empirical distribution, so that the
differences in the number of contacts for different models is quite
large.
%
Another interesting feature observed in this Figure, is that for the
eight models analyzed, there are \cetis{} that have no causal contacts
with any other \ceti{} during their whole lifetime.
%
The fraction of \cetis{} with no contact ranges from nearly 20\% up to
nearly 90\%, except for the dense and long lifetime models, for which
almost all \cetis{} make at least one contact.
%
Since these results depend on just eight models, we analyze the
properties of the frequency of contacts as a function of both the mean
lifetime and the mean awakening time.
%
\ffn{4}
%
On Fig.~\ref{F_never_contact} we show 2D color maps with the fraction
of \cetis{} in the simulations that never make contact (i.e., never
listen to another \ceti{}), as a function of the mean lifetime
($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
different values of the maximum signal range, $D_{max}=10000$~yr (left
panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
(right panel).
%
A clear pattern emerges in this Figure, showing that the probability
for a \ceti{} of making causal contact with at least another \ceti{}
during their entire lifetime, increases with increasing $D_{max}$,
increasing $\left<\tau_s\right>$ and decreasing $\left<\tau_a\right>$,
following a roughly linear dependence with the three parameters.

While the number of \cetis{} that do not succeed in reaching the
causal contact regions of other node is a useful indicator of
degree of isolation, there is also the chance that a number of
\cetis{} are already in the causal contact region of other nodes,
which shows a somewhat oposite situation.
%
\ffn{5}
%
The fraction of \cetis{} that make the first contact at the awakening
event is shown in the Fig.~\ref{F_C_at_A}, 
as a function of the mean lifetime
($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
different values of the maximum signal range, $D_{max}=10000$~yr (left
panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
(right panel).
%
These probabilities are consistent with the ones on the previous
Figure.
%
Although they are not complementary, it is clear that the dependence
with the three parameters is roughly linear in both cases.
%
DISCUTIR ESTOS RESUOLTADOS SEGUN LA LITERATURA

%

@@

In this subsection we analyze the distribution of the waiting times for
a first contact.
%
Such distribution can be considered to compute the probability for a
random \ceti{} to carry out a SETI project and spend a given time until
the first contact is made.

\ffn{6}
%
In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
waiting times for the first contact, for several models.
%
Upper panel (a) panel shows the histograms for several values of the
mean survival time $\tau_s$, and the mean awakening time $\tau_a$ in
the range 68000-100000 yr.
%
Lower panel (b) shows the histograms for several values of $\tau_a$,
and $\tau_s$ in the range 17000-24000 yr.
%
As it can be seen, the shape of the histograms is almost the same when
$\tau_a$ is modified (panel (b)), but significantly changes when
$\tau_s$ is modified (panel (a)). 
%
For different values of $\tau_a$, the shape of the distribution is
similar, and the main difference is given by the different number of
contacts on the different models.
%
Conversely, for different values of $\tau_a$, the shape of the
distribution is similar, and the main difference given by the
different number of contacts on the different models.

The fraction of the \cetis{} that have made at least one contact as a
function of the elapsed time since the awakening, is computed with
respect to the total number of nodes that make causal contact at least
one time in the time range from the Awakening up to the Doomsday
events.
%
From a frequentist approach, the total counts from the awakening
($t=$0~yr) up to a given time are related to the estimation of the
probability of listening during that time interval and reaching to
causal contact region of at least another node.
%
The complement of this value is the probability of observing in a time
interval with no success, i.e., without ever happening a Contact
event.
%
Clearly, this probability diminishes with time and tends to zero for
large time periods.
%

### 2020-01-08 manuscript.tex
acomodando las figuras
retocando las conclusiones

We have implemented a suite of simulations following a stochastic
approach, to explore the hypothesis space of a simple model that
accounts for the causal connections between communicating
civilizations on an simplified galaxy.
%
The different models can be generated with a minimal number of free
parameters and some fixed assumptions.
%
We argue that three parameters can be used to describe a variety of
situations, ranging from a Galaxy where an intelligent civilization is
very rare, to a Galaxy populated with plenty of civilizations in
causal contact.
%
We explored the consecuences of different models using mensurable
quantities, that arise as by-products of the simulations.
%
With this approach, we can estimate several quantities as a function
of the parameters on the hypothesis space.
%
For instance, it is possible to estimate the timescale for an
uninterrupted SETI effort to reach success, in terms of different
model parameters that reflect different, so far unknown, scenarios for
the spread of intelligent life on the galaxy.

Although the simulations use a number of speculative assumptions, we
argue that the degree of knowledge about the origin and persistence of
life in the Galaxy does not worth the implementation of more detailed
or sophisticated models.
%
Instead, we take advantadge of the simplicity of the model to explore
the hypothesis space, i.e., the three--dimensional parameter space of
the parametric model, in order to gain insight on the consecuences of
different scenarios for the search of intelligent life.
%
Our analysis is not centered in obtaining the odds for the Earth
making contact with another extraterrestrial civilization, but in
obtaining a statistical, parameter dependent description of the
communication system consisting of a group of broadcasting
civilizations which arise sparsely both in time and space in the
Galaxy.

@@

Taking into account the caveat mentioned previously about the
trade--off between the simplicity and complexity of the experiment,
future work can be performed following this framework in order to
explore possible implications of the results for more detailed
configurations.
%
For example, the communication method (isotropic, colimated,
serendipitous) can affect the observables.
%
Since we analyze the causal contacts, effective contacts depend on the
efficiency of both the emitter and the receiver.
%
A correction by coverage ratio in the detection and by a targetting
ratio in the emission could be easily implemented in the simulation,
although the effect would be to reduce the probability of contact by
an efficiency ratio.
%
Thus, such implementation is not really necessary, since the values of
probabilities could modified by a constant correction factor equal to
the combined emission/reception efficiency
\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
forgan_collimated_2014}.
%
If we consider a statistical distribution for this factor, then the
results could change, and this could be studied with the DE approach.
%
Also the nature of the message carrier, electromagnetic or another.
%
It is also possible to explore the effects of alignments or the use of
stars as sources or amplifiers.
%
%energetic markers Spiritual markers
%
As a future work, we will explore several improvements to the
simulation, and analyze the impact on the final results of these
changes.
%
They include a spatial distribution that actually resembles the shape
of the Galaxy, different distribution functions for the mean lifetime
of MPLs and different efficiencies for the sphere of the causal region
of each MPL, resembling different searching strategies
\citep{hippke_interstellar_2017}.

@@

\bibliographystyle{mn2e}
\bibliography{biblio_seti}

>>>>

We have implemented a suite of simulations following a stochastic
approach, to explore the hypothesis space of a simple model that
accounts for the causal connections between communicating
civilizations on a simplified galaxy.
%
The different models can be generated with a minimal number of free
parameters and some fixed assumptions.
%
We argue that three parameters can be used to describe a variety of
situations, ranging from a Galaxy where an intelligent civilization is
very rare, to a Galaxy populated with plenty of civilizations in
causal contact.
%
With this approach, we can estimate several quantities as a function
of the parameters on the hypothesis space,
so we explored the outcomes of different models 
that arise as by-products of the simulations.

%
For instance, it is possible to estimate the timescale for an
uninterrupted SETI effort to reach success, in terms of different
model parameters that reflect different, so far unknown, scenarios for
the appeareance of intelligent life on the galaxy.

Although the simulations use a number of speculative assumptions, we
argue that the degree of knowledge about the origin and persistence of
life in the Galaxy does not worth the implementation of more detailed
or sophisticated models.
%
Instead, we take advantadge of the simplicity of the model to explore
the hypothesis space, i.e., the three--dimensional parameter space of
the parametric model, in order to gain insight on the consecuences of
different scenarios for the search of intelligent life.
%
Our analysis is not centered in obtaining the odds for the Earth
to make contact with another intelligent civilization.
%
Instead, we focus on obtaining a statistical, parameter dependent
description of the possible properties of the communication networks
that comprise sets of nodes with broadcasting and reception
capabilities.
%
This causally connected nodes are sparsely distributed in both space
and time, making it difficult the analytical treatment and justifying
the simulation approach.

@@

Taking into account the caveat mentioned previously about the
trade--off between the simplicity and complexity of the experiment,
further analysis could be performed following this framework in order to
explore possible implications of the results for more detailed
configurations.
%
For example, the communication method (isotropic, colimated,
serendipitous) can affect the observables, making it neccesary to
implement a correction factor.
%
Taking this into account, our results regarding the probability of
causal contact should be considered as upper limits for effective
contacts, since they depend on the
efficiency of both the emitter and the receiver to broadcast and scan
the sky for intelligent signals, respectively.
%
A correction by coverage ratio in the detection and by a targetting
ratio in the emission could be easily implemented in the simulation,
although the effect of reducing the probability of contact
is basically the product of the efficiency rations for emmision and
detection and thus 
such implementation is not really necessary
and the values of
probabilities could modified by a constant correction factor equal to
the combined emission/reception efficiency
\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
forgan_collimated_2014}.
%
If we consider a statistical distribution for this factor, then the
results could change, and this could be studied with the DE approach.
%
Also the nature of the message carrier, electromagnetic or another.
%
It is also possible to explore the effects of alignments or the use of
stars as sources or amplifiers.
%
%energetic markers Spiritual markers
%
As a future work, we will explore several improvements to the
simulation, and analyze the impact on the final results of these
changes.
%
They include a spatial distribution that actually resembles the shape
of the Galaxy, different distribution functions for the mean lifetime
of MPLs and different efficiencies for the sphere of the causal region
of each MPL, resembling different searching strategies
\citep{hippke_interstellar_2017}.

@@

\bibliographystyle{mn2e}
\setlength{\bibsep}{0.0pt}

\bibliography{biblio_seti}
